---
title: "Email - Exploratory Analysis"
author: "Chloe Li"
date: "September 21, 2016"
output:
  pdf_document:
    highlight: monochrome
  html_document:
    highlight: pygments
    theme: united
  word_document: default
header-includes: \usepackage{graphicx} \usepackage{fancyhdr} \pagestyle{fancy} \setlength\headheight{28pt}
  \fancyhead[L]{\includegraphics[width=0.5cm]{mobovidalogo.png}} \fancyfoot[LE,RO]{CL}
classoption: landscape
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r load, message=FALSE, warning=FALSE, include=FALSE}

#--------------------------------------------------LOAD-------------------------------------------------------#
rm(list = ls())
#set directory
setwd("/Users/chloeli/Documents/02. EmailPrediction_CL/011.Data")

#load neccessary packages
#Note, you might need to install pacman first
pacman::p_load("gridExtra","ggplot2", "dplyr","lubridate","reshape2","data.table","quantmod","lme4","lattice","plyr","broom",'ReporteRs',"knitr","xtable", "cluster","fpc","factoextra","grid","gridExtra","scales")

      
#load data
RFM_DT <- read.csv("RRFFK_220K_sends.csv") #this is aggregated data after David Liu rank riid by recency and frequency

RFM_DT <- dplyr::filter(RFM_DT, sends != "#N/A")

#there are 3953 records of sends being "#N/A". David said remove those no matter what.


RFM_DT_orig <- RFM_DT #reassign the dataset into new vector for clustering purpose

#reformat data for graphing purpose
RFM_DT$riid <- as.character(RFM_DT$riid)
#change multiple variables into factor type
cols = c(2, 3, 4,5)
RFM_DT[cols] <- lapply(RFM_DT[cols], factor)

RFM_DT[,'sends'] <- as.numeric(as.character(RFM_DT[,'sends']))
RFM_DT[,'conversions'] <- as.numeric(RFM_DT[,'conversions'])
```

## Introduction
- This documents the exploratory analysis process for email prediction project.
- The sample of input data is shown as below restructured and sent by David Liu:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
kable(head(RFM_DT_orig))
```


## Explanations of variables
- riid: unique identifier of email receivers
- Ro: recency score of open. The higher the score is, the most recent action was taken by the customer.
- Rc: recency score of click. Same as Ro. Table below shows the number of days passed associated with the score.
- Fo: life time open rate; number of opens/number of sent. The higher the score is, the higher the rate is.
- Fc: life time click rate; number of clicks/number of sent. Same logic as above.
- conversions: number of times the customer converted.

```{r table_Var, echo=FALSE, message=FALSE, warning=FALSE, results='asis', warnings=FALSE}
tabl <- "
|Ro/Rc Score | 7 | 6 | 5  |  4 |  3 |  2 |   1     |
|-----------:|--:|--:|---:|---:|---:|---:|--------:|
|Days Before | 3 | 7 | 14 | 30 | 60 | 90 | LifeTime|  
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```


```{r graph_DT, message=FALSE, warning=FALSE, include=FALSE}

#------------------------------------------------TRANSFORM-----------------------------------------------------#
#this is a very stupid way of grouping multiple variables and summarise them

#this is grouping and summarising with sends

#transform the data from wide to long
#RFM_DT_gr <- melt(RFM_DT, id=c(1,6), measure=c("Ro", "Rc", "Fc","Fo"))

Groupby_Ro <- dplyr::group_by(RFM_DT, Ro)
SumRo <- dplyr::summarise(Groupby_Ro,
                          Total_conversions_Ro = sum(conversions),
                          Total_sends_Ro = sum(sends),
                          Num_riid_Ro = length(riid))

Groupby_Rc <- dplyr::group_by(RFM_DT, Rc)
SumRc <- dplyr::summarise(Groupby_Rc,
                          Total_conversions = sum(conversions),
                          Total_sends = sum(sends),
                          Num_riid = length(riid))


Groupby_Fc <- dplyr::group_by(RFM_DT, Fc)
SumFc <- dplyr::summarise(Groupby_Fc,
                          Total_conversions = sum(conversions),
                          Total_sends = sum(sends),
                          Num_riid = length(riid))

Groupby_Fo <- dplyr::group_by(RFM_DT, Fo)
SumFo <- dplyr::summarise(Groupby_Fo,
                          Total_conversions = sum(conversions),
                          Total_sends = sum(sends),
                          Num_riid = length(riid))


#combine dataset
SumRo$Total_conversions_Rc <- SumRc$Total_conversions
SumRo$Total_sends_Rc <- SumRc$Total_sends
SumRo$Num_riid_Rc <- SumRc$Num_riid



SumRo$Total_conversions_Fc <- SumFc$Total_conversions
SumRo$Total_sends_Fc <- SumFc$Total_sends
SumRo$Num_riid_Fc <- SumFc$Num_riid

SumRo$Total_conversions_Fo <- SumFo$Total_conversions
SumRo$Total_sends_Fo <- SumFo$Total_sends
SumRo$Num_riid_Fo <- SumFo$Num_riid



#assign the dataset into a new dataframe (just to change the name)
Summary_Membership_stat <- SumRo



#1. calculate the conversion/riid ratio
Summary_Membership_stat$converID_ratioRo <- Summary_Membership_stat$Total_conversions_Ro /Summary_Membership_stat$Num_riid_Ro


Summary_Membership_stat$converID_ratioRc <- Summary_Membership_stat$Total_conversions_Rc /Summary_Membership_stat$Num_riid_Rc


Summary_Membership_stat$converID_ratioFo <- Summary_Membership_stat$Total_conversions_Fo /Summary_Membership_stat$Num_riid_Fo

Summary_Membership_stat$converID_ratioFc <- Summary_Membership_stat$Total_conversions_Fc /Summary_Membership_stat$Num_riid_Fc



#2. calculate the conversion/sent

Summary_Membership_stat$converSent_ratioRo <- Summary_Membership_stat$Total_conversions_Ro /Summary_Membership_stat$Total_sends_Ro


Summary_Membership_stat$converSent_ratioRc <- Summary_Membership_stat$Total_conversions_Rc /Summary_Membership_stat$Total_sends_Rc


Summary_Membership_stat$converSent_ratioFo <- Summary_Membership_stat$Total_conversions_Fo /Summary_Membership_stat$Total_sends_Fo


Summary_Membership_stat$converSent_ratioFc <- Summary_Membership_stat$Total_conversions_Fc /Summary_Membership_stat$Total_sends_Fc



#3. sent/num of riid

Summary_Membership_stat$sentID_ratioRo <- Summary_Membership_stat$Total_sends_Ro /Summary_Membership_stat$Num_riid_Ro

Summary_Membership_stat$sentID_ratioRc <- Summary_Membership_stat$Total_sends_Rc /Summary_Membership_stat$Num_riid_Rc

Summary_Membership_stat$sentID_ratioFo <- Summary_Membership_stat$Total_sends_Fo /Summary_Membership_stat$Num_riid_Fo

Summary_Membership_stat$sentID_ratioFc <- Summary_Membership_stat$Total_sends_Fc /Summary_Membership_stat$Num_riid_Fc



Summary_Membership_stat <- dplyr::select(Summary_Membership_stat, RFM_Score = Ro, everything())
```





```{r ConverID, message=FALSE, warning=FALSE, include=FALSE}
#transform the data from wide to long
#on convert/number of riid ratio
ConvertID <- melt(Summary_Membership_stat, id=1, measure=c("converID_ratioRo", "converID_ratioRc", "converID_ratioFo","converID_ratioFc"))


#transform the data from wide to long
#on total of conversions
ConvertTotal <- melt(Summary_Membership_stat, id=1, measure=c("Total_conversions_Ro", "Total_conversions_Rc", "Total_conversions_Fo","Total_conversions_Fc"))

```

\pagebreak

# Comparison Graphs

- Total Conversions

```{r ConverID_g01, echo=FALSE, message=FALSE, warning=FALSE}
#graph a scatter plot for each RFM score and see how much conversions/number of riid relate to each categories
ggplot(ConvertTotal, aes(x=RFM_Score, y=value, colour=variable, shape=variable)) + 
  geom_point(size = 4) + 
  scale_size_area() + 
  #scale_colour_brewer(palette = "Set1") +
  ggtitle("Total Conversions") +
  ylab("Total Conversions") +
  theme(legend.position ="top", 
  panel.background = element_rect(fill="light grey"))  + scale_shape_discrete(labels=c("Ro","Rc","Fo","Fc")) + scale_colour_discrete(labels=c("Ro","Rc","Fo","Fc"))
  


```

- Conversions/Number of riid

```{r ConverID_g02, echo=FALSE, message=FALSE, warning=FALSE}
#graph a scatter plot for each RFM score and see how much conversions/number of riid relate to each categories
ggplot(ConvertID, aes(x=RFM_Score, y=value, colour=variable, shape=variable)) + 
  geom_point(alpha=1,size = 4) + 
  scale_size_area() + 
  #scale_colour_brewer(palette = "Set1") +
  ggtitle("Conversions per Membership") +
  ylab("Conversions per Membership") +
  theme(legend.position ="top",  
  panel.background = element_rect(fill="light grey")
) + scale_y_continuous(labels = scales::percent)



#grid.arrange(a, b, legend, ncol=2, #, nrow = 2)
#             layout_matrix = rbind(c(1,2), c(3,3)))
            # widths = c(2.7, 2.7), heights = c(2.5, 0.2))



#source("/Users/chloeli/Documents/02. EmailPrediction_CL/017. R_Function_CL_Email/multiplot_fun.R")

#multiplot(a, b, cols=2)


#http://www.sthda.com/english/wiki/ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page-r-software-and-data-visualization

```

\pagebreak

```{r SentID, message=FALSE, warning=FALSE, include=FALSE}
#transform the data from wide to long
#on sents/number of riid ratio
SentID <- melt(Summary_Membership_stat, id=1, measure=c("sentID_ratioRo", "sentID_ratioRc", "sentID_ratioFo","sentID_ratioFc"))


#transform the data from wide to long
#on total of sents
SentTotal <- melt(Summary_Membership_stat, id=1, measure=c("Total_sends_Ro", "Total_sends_Rc", "Total_sends_Fo","Total_sends_Fc"))

```


## Total Sends

```{r sent_g01, echo=FALSE, message=FALSE, warning=FALSE}
#graph a scatter plot for each RFM score and see how much conversions/number of riid relate to each categories
#options(scipen=10000)

ggplot(SentTotal, aes(x=RFM_Score, y=value, colour=variable, shape=variable)) + 
  geom_point(alpha=1,size = 4) + 
  scale_size_area() + 
  scale_colour_brewer(palette = "Set1") +
  ggtitle("Total Sents") +
  ylab("Total Sents") + theme(legend.position ="top", 
  panel.background = element_rect(fill="light grey")) + scale_y_continuous(labels = comma) + scale_shape_discrete(labels=c("Ro","Rc","Fo","Fc")) + scale_colour_discrete(labels=c("Ro","Rc","Fo","Fc"))

```

\pagebreak

## Sent by membership

```{r sent_g02, echo=FALSE, message=FALSE, warning=FALSE}
#graph a scatter plot for each RFM score and see how much conversions/number of riid relate to each categories
ggplot(SentID, aes(x=RFM_Score, y=value, colour=variable, shape=variable)) + 
  geom_point(alpha=1,size = 4) + 
  scale_size_area() + 
  scale_colour_brewer(palette = "Set1") +
  ggtitle("Sent per membership") +
  ylab("Sent per membership") + theme(legend.position ="top", 
  panel.background = element_rect(fill="light grey")) + scale_shape_discrete(labels=c("Ro","Rc","Fo","Fc")) + scale_colour_discrete(labels=c("Ro","Rc","Fo","Fc"))

```

\pagebreak


```{r ConverSent, message=FALSE, warning=FALSE, include=FALSE}
#transform the data from wide to long
#on conversions/sent
ConverSent <- melt(Summary_Membership_stat, id=1, measure=c("converSent_ratioRo", "converSent_ratioRc", "converSent_ratioFo","converSent_ratioFc"))

```

## Conversion per sent

```{r sent_g03, echo=FALSE, message=FALSE, warning=FALSE}
#graph a scatter plot for each RFM score and see how much conversions/number of riid relate to each categories
ggplot(ConverSent, aes(x=RFM_Score, y=value, colour=variable, shape=variable)) + 
  geom_point(alpha=1,size = 4) + 
  scale_size_area() + 
  scale_colour_brewer(palette = "Set1") +
  ggtitle("Conversion per Sent") +
  ylab("Conversion per Sent") + theme(legend.position ="top", 
  panel.background = element_rect(fill="light grey")
) +scale_y_continuous(labels = scales::percent)+ scale_shape_discrete(labels=c("Ro","Rc","Fo","Fc")) + scale_colour_discrete(labels=c("Ro","Rc","Fo","Fc")) 

```



\pagebreak

## Histogram on four groups to count number of riid per score.
- To see how many membership per score for each group.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#stupid way to graph histograms in one plot
PRo = ggplot(RFM_DT_orig, aes(x=Ro)) + 
  geom_histogram(bins = 30,fill = "#CC6666") + ylab("Number of Membership") + theme(
  panel.background = element_rect(fill="light grey")
)
PRc = ggplot(RFM_DT_orig, aes(x=Rc)) + 
  geom_histogram(bins = 30,fill = "#0072B2") + ylab("Number of Membership") + theme(
  panel.background = element_rect(fill="light grey")
)
PFc = ggplot(RFM_DT_orig, aes(x=Fc)) + 
  geom_histogram(bins = 30,fill = "#009E73") + ylab("Number of Membership") + theme(
  panel.background = element_rect(fill="light grey")
)
PFo = ggplot(RFM_DT_orig, aes(x=Fo)) + 
  geom_histogram(bins = 30,fill = "#9999CC") + ylab("Number of Membership") + theme(
  panel.background = element_rect(fill="light grey")
)

source("/Users/chloeli/Documents/02. EmailPrediction_CL/017. R_Function_CL_Email/multiplot_fun.R")

multiplot(PRo,PRc,PFc, PFo, cols=2)


```


```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(ggplot2)
p1 = qplot(1:10,rnorm(10))
p2 = qplot(1:10,rnorm(10))
library(gridExtra)
grid.arrange(p1, p2, ncol=2, top = "Main title")

```

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Determine number of clusters

k_ready <- RFM_DT_orig[,-1]
k_ready <- k_ready[,-5]

#wss <- (nrow(k_ready)-1)*sum(apply(k_ready,2,var))
#for (i in 2:15) wss[i] <- sum(kmeans(k_ready, 
#  	centers=i)$withinss)
#plot(1:15, wss, type="b", xlab="Number of Clusters",
#  ylab="Within groups sum of squares")

fviz_nbclust(k_ready, kmeans, method = "wss") +
    geom_vline(xintercept = 3, linetype = 2)


# K-Means Cluster Analysis
fit <- kmeans(k_ready, 4) # 4 cluster solution
# get cluster means 
aggregate(k_ready,by=list(fit$cluster),FUN=mean)
# append cluster assignment
RFM_DT_orig <- data.frame(RFM_DT_orig, fit$cluster)
```


```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#set directory
setwd("/Users/chloeli/Documents/02. EmailPrediction_CL/011.Data")

email_users <- read.csv("emails_useremailstats_partial.csv")




```


